---
title: "Lead Litigations"
author: "Reese Hirota"
output: html_document
date: "2023-04-11"
---

Maps and charts on lead litigations.

```{r setup, include=FALSE}
# ---- SETUP ----
setwd("/Users/rhirota/Documents/GitHub/lead_hearing")
source("code/utils/00_load_dependencies.R")
```

Data Source:
https://data.cityofnewyork.us/Housing-Development/Housing-Litigations/59kj-x8nc

```{r}
# ---- LOAD DATA ----
litigations <-fread("https://data.cityofnewyork.us/resource/59kj-x8nc.csv?$limit=99999999999")
litigations[, yr := year(caseopendate)]
litigations <- litigations[, yr := year(caseopendate)]
litigations <- litigations[litigations$yr>2005,] # not complete data
```

```{r}
# ---- CLEAN DATA ----
lead_litigations <- unique(litigations[grep("lead", casetype, ignore.case = TRUE), ])
lead_litigations <- unique(lead_litigations[!grep("non-lead", casetype, ignore.case = TRUE), ])
lead_litigations[, lit_bbl := length(unique(litigationid)), by = .(bbl)]
```

```{r}
# ---- BASIC GGPLOTS ----
# ---- cleanup/setup for prp/pct plots
temp <- lead_litigations %>% group_by(yr) %>% summarise(count_lead=n())
temp2 <- litigations %>% group_by(yr) %>% summarise(count_total=n())
temp <- temp2 %>% full_join(y = temp, by = 'yr')
temp <- temp[temp$yr <= 2023 & !is.na(temp$yr) & !is.na(temp$count_lead), ] # not sure why NA/2030 is in there... data entry error?

# ---- plot number of lead litigations by year
lead_plot <- ggplot(data=(lead_litigations %>% group_by(yr) %>% summarise(count=n())), aes(x=yr, y=count)) + geom_line() + labs(title="Lead Litigation Cases (August 2006-Present)", subtitle="Data Provided by HPD", x="Year Case Opened", y="Number of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_plot, filename = "visuals/april_2023/lead_litigations.png", width = 8, height = 5)

# ---- plot number of lead litigations by year, by casetype
lead_casetype_plot <- ggplot(data=(lead_litigations %>% group_by(yr, casetype) %>% summarise(count=n())), aes(x=yr, y=count, col=casetype)) + geom_line() + labs(title="Lead Litigation Cases by Casetype (August 2006-Present)",subtitle="Data Provided by HPD", x="Year Case Opened", y="Number of Cases", color="Case Type") + theme_nycc() + scale_color_nycc()
ggsave(lead_casetype_plot, filename = "visuals/april_2023/lead_litigations_by_casetype.png", width = 8, height = 5)

# ---- plot prop lead litigations/total litigations by year
lead_prp <- ggplot(data=temp, aes(x=yr, y=count_lead/count_total)) + geom_line() + labs(title="Lead Litigations: Proportion of Total Cases (August 2006-Present)", subtitle="Data Provided by HPD", x="Year Case Opened", y="Proportion of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_prp, filename = "visuals/april_2023/lead_litigations_prp_total.png", width = 8, height = 5)
# ---- format above plot as pct
lead_pct <- ggplot(data=temp, aes(x=yr, y=count_lead/count_total)) + scale_y_continuous(labels=scales::percent) + geom_line() + labs(title="Lead Litigations: Percent of Total Cases (August 2006-Present)",subtitle="Data Provided by HPD", x="Year Case Opened", y="Percent of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_pct, filename = "visuals/april_2023/lead_litigations_pct_total2.png", width = 8, height = 5)
```

xxx

Possible interest: Open judgement (yes) or harassment found or penalty assigned

Open Judgement: A “Y” indicates that civil penalties are owed to HPD after the Court issued a judgment against the owners of a building awarding civil penalties to HPD.

Finding of Harassment

After Inquest: Judgments entered for an order and/or civil penalties upon default where there was no appearance in Court by the named respondent(s). If any of these judgments are vacated by the Court, they will be removed from this list.

After Trial: Judgments awarded by the Court at the conclusion of trial at which evidence was presented.

xxx

```{r}
council_districts <- council_districts %>%
  rename("council_district" = "CounDist")
lead_lit_cd <- lead_litigations %>% full_join(y = council_districts, by = 'council_district') %>%
                st_as_sf() 
lead_lit_cd <- lead_lit_cd[lead_lit_cd$council_district != 51,] # removing this one because there's no obs. in cd 51...
# ---- look at particularly bad litigation cases?
lead_lit_cd <- lead_lit_cd %>% mutate(findingofharassment = na_if(findingofharassment, ""))
judge_harass_fine <- filter(lead_lit_cd,
                            grepl("After", findingofharassment, fixed = TRUE) | 
                              casejudgement == "YES" |
                              !is.na(penalty))
# only 4 obs. where conditions met... lets see if there are any for all litigations
temp <- litigations %>% mutate(findingofharassment = na_if(findingofharassment, ""))
temp <- filter(litigations,
                            grepl("After", findingofharassment, fixed = TRUE) | 
                              casejudgement == "YES" |
                              !is.na(penalty)) # 4629 (out of 193927 obs total), 2.4%
                              # compared to >.001% of lead litigations; so less common
                              # for lead litigations but not super common regardless
```

Find lead litigations per thousand buildings by borough...

```{r}
buildings_in_boro <- pluto_sub %>% group_by(borough) %>% summarise(total_buildings=n())
lead_lit_boro <- lead_litigations %>% group_by(boroid) %>% summarise(num_lit=n())
```

```{r}
buildings_in_boro$boroid <- buildings_in_boro$borough
buildings_in_boro <- buildings_in_boro %>% mutate(boroid=recode(boroid, "MN" = 1,
                                           "BX" = 2,
                                           "BK" = 3,
                                           "QN" = 4,
                                           "SI" = 5))
```

```{r}
boro_redo <- buildings_in_boro %>% full_join(y = lead_lit_boro, by = 'boroid')
boro_redo$lit_per_thousand_buildings <- boro_redo$num_lit/(boro_redo$total_buildings/1000)
```

```{r}
# ---- write out csv of boro data
write.csv(boro_redo, "data/output/litgtns_per_thousand_bldngs_by_boro.csv", row.names=FALSE)
```

```{r}
boro_redo$units <- "added back in just for formatting"
boro_redo$per_thousand_units <- "added added back in just for formatting"
```

```{r}
# new format: boro / units / buildings/ count / per thousand / per thousand units
boro_redo_formatted <- boro_redo[,c(1, 6, 2, 4, 5, 7)]
boro_redo_formatted <- boro_redo_formatted %>% 
    rename("boro" = "borough",
           "bldngs" = "total_buildings",
           "count" = "num_lit",
           "per_thousand_bldngs" = "lit_per_thousand_buildings") %>%
  mutate(boro=recode(boro, "MN" = "MANHATTAN", 
                     "BX" = "BRONX",
                     "BK" = "BROOKLYN",
                     "QN" = "QUEENS",
                     "SI" = "STATEN ISLAND"))
# write.csv(boro_redo_formatted, "data/output/litgtns_per_thousand_bldngs_by_boro.csv", row.names=FALSE)
```
```{r}
# ---- plot boro data
temp <- ggplot(data=boro_redo, aes(x=borough, y=lit_per_thousand_buildings)) + geom_bar(stat="identity") + theme_nycc() + scale_color_nycc()
ggsave(temp, filename = "visuals/april_2023/litgtns_per_thousand_bldngs_by_boro.png", width = 8, height = 5)
```

```{r}
bar <- 
  ggplot(data=boro_redo_formatted,
         aes(reorder(x=boro,per_thousand_bldngs),
             y=per_thousand_bldngs,
             tooltip = paste0(boro, "\n",
                          round(per_thousand_bldngs,2),
                          " HPD Lead Litigations"))) +
  coord_flip() +
  geom_col_interactive(width = 0.8, fill = "#2F56A6") +
  theme_nycc() +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(y="HPD Lead Paint Litigations per 1000 buildings",
       x=element_blank())

bar

tooltip_css = "background-color:#CACACA;"

plot_interactive = girafe(ggobj = bar,  
                          width_svg = 8,
                          height_svg = 5,
                          options = list(opts_tooltip(css = tooltip_css)))

save_html(plot_interactive, file.path("visuals", "april_2023", "interactive_per_thousand_buildings_by_boro.html") )
```
