---
title: "Lead Litigations"
author: "Reese Hirota"
output: html_document
date: "2023-04-11"
---

Maps and charts on lead litigations.

```{r setup, include=FALSE}
# ---- SETUP ----
knitr::opts_chunk$set(echo = TRUE)
source("../../../code/utils/bin_geocode.R")
source("../../../code/utils/utils.R")
source("../../../code/utils/00_load_dependencies.R")
```

```{r}
# ---- SETUP ----
# ---- function to unzip shp
unzip_sf <- function(zip_url) {
  temp <- tempfile()
  temp2 <- tempfile()
  #download the zip folder from the internet save to 'temp' 
  download.file(zip_url, temp)
  #unzip the contents in 'temp' and save unzipped content in 'temp2'
  unzip(zipfile = temp, exdir = temp2)
  #if returns "character(0), then .shp may be nested within the folder
  your_SHP_file <- ifelse(!identical(list.files(temp2, pattern = ".shp$",full.names=TRUE), character(0)), 
                          list.files(temp2, pattern = ".shp$",full.names=TRUE), 
                          list.files(list.files(temp2, full.names=TRUE), pattern = ".shp$", full.names = TRUE))
  unlist(temp)
  unlist(temp2)
  return(your_SHP_file)
}
```

```{r}
# ---- SETUP ----
# ---- unzip cd shp
council_districts = unzip_sf("https://www.nyc.gov/assets/planning/download/zip/data-maps/open-data/nycc_21d.zip") %>%
  st_read() %>%
  st_transform(st_crs(4326))
```

Data Source:
https://data.cityofnewyork.us/Housing-Development/Housing-Litigations/59kj-x8nc

```{r}
# ---- SETUP ----
# ---- get bbls per council district
library(data.table)

pluto_file <- fread("https://data.cityofnewyork.us/resource/64uk-42ks.csv?$limit=99999999") 

pluto_sub <- pluto_file[unitsres > 0, .(
    bbl = str_sub(as.character(bbl), 1, 10),
    borough = borough,
    council_dist = council)]
```


```{r}
# ---- LOAD DATA ----
litigations <-fread("https://data.cityofnewyork.us/resource/59kj-x8nc.csv?$limit=99999999999")
litigations[, yr := year(caseopendate)]
litigations <- litigations[, yr := year(caseopendate)]
litigations <- litigations[litigations$yr>2005,] # not complete data
```

```{r}
# ---- CLEAN DATA ----
lead_litigations <- unique(litigations[grep("lead", casetype, ignore.case = TRUE), ])
lead_litigations <- unique(lead_litigations[!grep("non-lead", casetype, ignore.case = TRUE), ])
# lead_litigations[, caseopendate := as.Date(caseopendate)] # added above instead
# lead_litigations[, yr := year(caseopendate)]
lead_litigations[, lit_bbl := length(unique(litigationid)), by = .(bbl)]
```

```{r}
# ---- BASIC GGPLOTS ----
# ---- cleanup/setup for prp/pct plots
# litigations <- litigations[, yr := year(caseopendate)] # should've added this in prev section oops
temp <- lead_litigations %>% group_by(yr) %>% summarise(count_lead=n())
temp2 <- litigations %>% group_by(yr) %>% summarise(count_total=n())
temp <- temp2 %>% full_join(y = temp, by = 'yr')
temp <- temp[temp$yr <= 2023 & !is.na(temp$yr) & !is.na(temp$count_lead), ] # not sure why NA/2030 is in there... data entry error?

# ---- plot number of lead litigations by year
lead_plot <- ggplot(data=(lead_litigations %>% group_by(yr) %>% summarise(count=n())), aes(x=yr, y=count)) + geom_line() + labs(title="Lead Litigation Cases (August 2006-Present)", subtitle="Data Provided by HPD", x="Year Case Opened", y="Number of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_plot, filename = "../../../visuals/lead_hearing_4_25_23/lead_litigations.png", width = 8, height = 5)

# ---- plot number of lead litigations by year, by casetype
lead_casetype_plot <- ggplot(data=(lead_litigations %>% group_by(yr, casetype) %>% summarise(count=n())), aes(x=yr, y=count, col=casetype)) + geom_line() + labs(title="Lead Litigation Cases by Casetype (August 2006-Present)",subtitle="Data Provided by HPD", x="Year Case Opened", y="Number of Cases", color="Case Type") + theme_nycc() + scale_color_nycc()
ggsave(lead_casetype_plot, filename = "../../../visuals/lead_hearing_4_25_23/lead_litigations_by_casetype.png", width = 8, height = 5)

# ---- plot prop lead litigations/total litigations by year
lead_prp <- ggplot(data=temp, aes(x=yr, y=count_lead/count_total)) + geom_line() + labs(title="Lead Litigations: Proportion of Total Cases (August 2006-Present)", subtitle="Data Provided by HPD", x="Year Case Opened", y="Proportion of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_prp, filename = "../../../visuals/lead_hearing_4_25_23/lead_litigations_prp_total.png", width = 8, height = 5)
# ---- format above plot as pct
lead_pct <- ggplot(data=temp, aes(x=yr, y=count_lead/count_total)) + scale_y_continuous(labels=scales::percent) + geom_line() + labs(title="Lead Litigations: Percent of Total Cases (August 2006-Present)",subtitle="Data Provided by HPD", x="Year Case Opened", y="Percent of Cases") + theme_nycc() + scale_color_nycc()
ggsave(lead_pct, filename = "../../../visuals/lead_hearing_4_25_23/lead_litigations_pct_total.png", width = 8, height = 5)

# lead_num_vs_pct <- grid.arrange(lead_plot, lead_pct, nrow = 1)
# ggsave(lead_num_vs_pct, filename = "../../../visuals/litigations_num_vs_pct.png", width = 14, height = 5) # pretty similar
```

```{r}
# ---- MAP 1 ----
# ---- merge lead data w/ council district shp
council_districts <- council_districts %>%
  rename("council_district" = "CounDist")
lead_lit_cd <- lead_litigations %>% full_join(y = council_districts, by = 'council_district') %>%
                st_as_sf() 
lead_lit_cd <- lead_lit_cd[lead_lit_cd$council_district != 51,] # removing this one because there's no obs. in cd 51...
```

```{r}
# ---- MAP 1 ----
# ---- group sums, set up color palette
lead_lit_cd_sums <- lead_lit_cd %>% group_by(geometry, council_district) %>% summarise(count=n())

pal = colorNumeric(
  palette = "Blues",
  domain = lead_lit_cd_sums$count, 
  na.color = "White", 
  reverse = FALSE
)
```

Next map: merge w/ pluto, normalize by number of buildings (litigations/number of buildings)
merge pluto w/ CD

```{r}
# ---- MAP 1 ----
# ---- all lead litigations on opendata, compiled since august 2006
all_leadlit <- leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 11, maxZoom = 16)) %>% 
  addCouncilStyle(add_dists = TRUE) %>%
  addPolygons(data=lead_lit_cd_sums, weight=1, col=~pal(count), fillOpacity = .7, 
              popup = paste0("<strong>Number of litigations: </strong>", lead_lit_cd_sums$count, "<br>")) %>%
  addLegend("topleft", pal = pal, values = unique(lead_lit_cd_sums$count), title = "Lead Litigations: 2006-Present", opacity = 1, position = "topleft")
# all_leadlit
```

```{r}
# ---- MAP 1 ----
# ---- save map 1: html
saveWidget(all_leadlit, file=file.path('../../../visuals/lead_hearing_4_25_23', 
                               "lead_litigations_by_cd.html"))

# ---- save map 1: png
mapview::mapshot(all_leadlit, 
        file = file.path('../../../visuals/lead_hearing_4_25_23', "lead_litigations_by_cd.png"),
        remove_controls = c("homeButton", "layersControl"), vwidth = 1000, vheight = 850)
```

xxx

Possible interest: Open judgement (yes) or harassment found or penalty assigned

Open Judgement: A “Y” indicates that civil penalties are owed to HPD after the Court issued a judgment against the owners of a building awarding civil penalties to HPD.

Finding of Harassment

After Inquest: Judgments entered for an order and/or civil penalties upon default where there was no appearance in Court by the named respondent(s). If any of these judgments are vacated by the Court, they will be removed from this list.

After Trial: Judgments awarded by the Court at the conclusion of trial at which evidence was presented.

xxx

```{r}
# ---- look at particularly bad litigation cases?
lead_lit_cd <- lead_lit_cd %>% mutate(findingofharassment = na_if(findingofharassment, ""))
judge_harass_fine <- filter(lead_lit_cd,
                            grepl("After", findingofharassment, fixed = TRUE) | 
                              casejudgement == "YES" |
                              !is.na(penalty))
# only 4 obs. where conditions met... lets see if there are any for all litigations
temp <- litigations %>% mutate(findingofharassment = na_if(findingofharassment, ""))
temp <- filter(litigations,
                            grepl("After", findingofharassment, fixed = TRUE) | 
                              casejudgement == "YES" |
                              !is.na(penalty)) # 4629 (out of 193927 obs total), 2.4%
                              # compared to >.001% of lead litigations; so less common
                              # for lead litigations but not super common regardless
```

```{r}
# ---- MAP 2 ----
# ---- circle map format and palette
circle_map <- distinct(lead_litigations) %>%
  mutate(addr = paste(housenumber, streetname))

pal2 = colorNumeric(
  palette = "Blues",
  domain = circle_map$lit_bbl,
  reverse = FALSE
)
```
 
```{r}
# ---- MAP 2 ----
# ---- all lead litigations on opendata, compiled since august 2006
all_leadlit_circlevers <- leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 11, maxZoom = 16)) %>% 
  addPolygons(data=council_districts, weight=1, col='#d3d3d3', fillOpacity = 0.2) %>%
  addCouncilStyle(add_dists = TRUE) %>%
  addCircles(data=circle_map,
                 lng=~longitude, lat=~latitude,
                 color = ~pal2(lit_bbl), 
                 popup = paste0("<strong>Number of litigations: </strong>", circle_map$lit_bbl, "<br>",
                                "<strong>Address: </strong>", circle_map$addr, "<br>",
                                "<strong>Respondent: </strong>", circle_map$respondent, "<br>"),
                 group = df,
                 label =  ~paste("Lead Litigations since August 2006:", circle_map$lit_bbl),
                 weight = 2,
                 radius = 1, 
                 stroke = 2, 
                 opacity = .5, 
                 labelOptions = labelOptions(noHide = F,
                 direction = 'auto')) %>% 
   addLegend("topleft", pal = pal2, values = unique(circle_map$lit_bbl), title = "Lead Litigations: 2006-Present", opacity = 1, position = "topleft", bins=4)
# all_leadlit_circlevers
```

```{r}
# ---- MAP 2 ----
# ---- save map 2: html
saveWidget(all_leadlit_circlevers, file=file.path('../../../visuals/lead_hearing_4_25_23', 
                               "lead_litigations_by_bbl.html"))

# ---- save map 2: png
mapview::mapshot(all_leadlit_circlevers, 
        file = file.path('../../../visuals/lead_hearing_4_25_23', "lead_litigations_by_bbl.png"),
        remove_controls = c("homeButton", "layersControl"), vwidth = 1000, vheight = 850)
```

```{r}
# ---- MAP 1.5 ----
# ---- merge pluto data with council district shp
pluto_sub <- pluto_sub %>% rename("council_district" = "council_dist")
pluto_sub_cd <- pluto_sub %>% full_join(y = council_districts, by = 'council_district') %>% st_as_sf() 

# ----  get prop of lead litigations per thousand buildings
pluto_sub_cd_sums <- pluto_sub_cd %>% group_by(geometry, council_district) %>% summarise(count=n())
lit_prop <- left_join(lead_lit_cd_sums, pluto_sub_cd_sums %>% as.data.frame() %>% select(-geometry), by = "council_district")
lit_prop$lit_per_thous_bldngs <- (lit_prop$`count.x`/lit_prop$`count.y`)*1000
lit_prop <- lit_prop %>% select('geometry','council_district', lit_per_thous_bldngs)
```

```{r}
# ---- MAP 1.5 ----
# ---- set up color palette
pal = colorNumeric(
  palette = "Blues",
  domain = lit_prop$lit_per_thous_bldngs, 
  na.color = "White", 
  reverse = FALSE
)
```

```{r}
# ---- MAP 1.5 ----
lit_per_thou <- leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 11, maxZoom = 16)) %>% 
  addCouncilStyle(add_dists = TRUE) %>%
  addPolygons(data=lit_prop, weight=1, col=~pal(lit_per_thous_bldngs), fillOpacity = .7,
              popup = paste0("<strong>Number of litigations per thousand buildings: </strong>", round(lit_prop$lit_per_thous_bldngs,2), "<br>")) %>%
  addLegend("topleft", pal=pal, values=unique(lit_prop$lit_per_thous_bldngs), title="Lead Litigations (per Thousand Buildings): 2006-Present", opacity = 1, position = "topleft")
```

```{r}
# ---- MAP 1.5 ----
# ---- save map 1.5: html
saveWidget(lit_per_thou, file=file.path('../../../visuals/lead_hearing_4_25_23', 
                               "lead_litigations_per_thsnd_bbl_by_cd.html"))

# ---- save map 1.5: png
mapview::mapshot(lit_per_thou, 
        file = file.path('../../../visuals/lead_hearing_4_25_23',
                         "lead_litigations_per_thsnd_bbl_by_cd.png"),
        remove_controls = c("homeButton", "layersControl"), vwidth = 1000, vheight = 850)
```


```{r}
buildings_in_boro <- pluto_sub %>% group_by(borough) %>% summarise(total_buildings=n())
```

```{r}
lead_lit_boro <- lead_litigations %>% group_by(boroid) %>% summarise(num_lit=n())
```

```{r}
# df %>% mutate(result=recode(result, 'Win'='1', .default=NA_character_))
buildings_in_boro <- buildings_in_boro %>% mutate(boroid=recode(boroid, "MN" = 1,
                                           "BX" = 2,
                                           "BK" = 3,
                                           "QN" = 4,
                                           "SI" = 5))
```


```{r}
boro_redo <- buildings_in_boro %>% full_join(y = lead_lit_boro, by = 'boroid')
```

```{r}
boro_redo$lit_per_thousand_buildings <- boro_redo$num_lit/(boro_redo$total_buildings/1000)
```

```{r}
boro_redo
write.csv(boro_redo, "../../../data/litgtns_per_thousand_bldngs_by_boro.csv", row.names=FALSE)
```
```{r}
temp <- ggplot(data=boro_redo, aes(x=borough, y=lit_per_thousand_buildings)) +
  geom_bar(stat="identity")
ggsave(temp, filename = "../../../visuals/lead_hearing_4_25_23/litgtns_per_thousand_bldngs_by_boro.png", width = 8, height = 5)
```

